# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
Read the comma seperated file
```{r}
data<-read.csv("activity.csv")
```

Convert the date column from factor to date.
```{r}
data<-transform(data,date= as.Date(date,format="%Y-%m-%d"))
```

## What is mean total number of steps taken per day?
Count the steps for each day
```{r}
steps.per.day<-tapply(data$steps,data$date,sum,na.rm=TRUE)
barplot(steps.per.day,main="Total Steps Per Day",ylab="Steps",xlab="Date")
```
Mean steps per day
```{r}
mean(steps.per.day)
```
Median steps per day
```{r}
median(steps.per.day)
```




## What is the average daily activity pattern?
Count the steps for each 5 minute interval 
```{r}
steps.per.interval<-tapply(data$steps,data$interval,mean,na.rm=TRUE)
```
Plot the average steps per interval
```{r}
plot(steps.per.interval,
     type="l",main="Average Steps Per Interval",
     xlab="Interval",ylab="Steps")
```

The 5-minute interval, on average across all the days with the maximum number of steps
```{r}
which.max(steps.per.interval)
```

## Imputing missing values
The total number of rows with missing values
```{r}
sum(!complete.cases(data))
```
Fill in the missing values with average for that invterval
```{r}
clean.data<-data
for(i in 1:nrow(data)){
    if(is.na(data$steps[i]))
    { 
        clean.data$steps[i]<-steps.per.interval[as.character(data$interval[i])]
    }
}
```

Count the steps for each day
```{r}
clean.steps.per.day<-tapply(clean.data$steps,clean.data$date,sum)
barplot(clean.steps.per.day,main="Total Steps Per Day",sub="With imputed missing values",
        ylab="Steps",xlab="Date")
```
Mean steps per day
```{r}
mean(clean.steps.per.day)
```
Median steps per day
```{r}
median(clean.steps.per.day)
```
As you can see the mean and median were affected.  
## Are there differences in activity patterns between weekdays and weekends?
Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day.
```{r}
weekday<-factor(levels=c("weekday","weekend"))
for(i in 1:nrow(data)){
    if(weekdays(data$date[i]) %in% c("Monday","Tuesday","Wednesday","Thursday","Friday"))
        weekday[i]<-"weekday"
    else weekday[i]<-"weekend"
}
wk.data<-cbind(data,weekday)
```
A panel plot containing the 5-minute interval and the average number of steps taken, averaged across all weekday days or weekend days.
```{r}
library(lattice)
wk.data.avg<-aggregate(steps ~ interval + weekday, data=wk.data,FUN = mean)
xyplot(steps~interval|weekday,data=wk.data.avg,type="l")
```



